Spin canonical divisors of log stacky curves

### 2015-06-25 stacky_write_up/spin_cring.tex
fixed some spaced-out refs, accent grobner

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$ then 
$R(\sx,\Delta, \halfcan)$ has minimal relations in degree at most $\max(11, 2e_1,
\ldots, 2e_r).$ Otherwise, $R(\sx, \Delta, \halfcan)$ has minimal relations in degree at most
$\max(10, 2e_1, \ldots, 2e_r),$ so long as $\sigma$ does not lie in a finite list of
exceptional cases, as listed in Table ~\ref{table:g - 1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g_0_exceptional}
for signatures with $g = 0$.
\end{thm}

@@

Note that in the course of the proof of Theorem
~\ref{thm:g_0_generators_relations} we will not merely obtain
bounds on the degrees of the generators and relations. In fact, if
the genus $g = 0$ or $g = 1$, the proof yields an explicit system
of generators and the initial ideal of relations. While in the
genus $g \geq 2$ case, the proof does not always yield an explicit
system of generators and relations, if the halfcanonical divisor
$\halfcan$ can be written as $\halfcan' + \Lambda$ with $(\halfcan',
0, X)$ a log spin curve, and a system of generators and relations for
$ \halfcan'$ is known, then our proof provides an inductive
procedure for determining the generators and initial ideal of
relations for $ \halfcan$. In fact, such a system of generators and
relations are known for $\halfcan'$ in many cases, as detailed in
Neves ~\cite[ Section III.4]{neves:halfcan}.

@@

\section{Background}
\label{sec:background}
\todo{Reference independence of base field}
Look at Spencer-Kodaira \cite{kodaira:complex-manifolds}.
Include the following:
\begin{enumerate}
	\item Note that this result is independence of base field
	\item background on grobner bases
	\item Notation $h^0 = \dim H^0$.
	\item Why we have these fractions $\frac{e_i - 1}{2e_i}$ and why $e_i$ must be odd.
\end{enumerate}

@@

\noindent
determines map of rings, no longer necessarily graded. 
By assumption, $I'$ is generated over $I$ as a $k$-vector space by elements $f_1, \ldots, f_n$ whose initial terms have degree 2 and are linearly independent. By the theory of Grobner basis,
\[
	\dim_k(B'/B)_2=\dim_k(k[x_1, \ldots, x_n, y_1, \ldots, y_m]/k[x_1, \ldots, x_n])_2 - j
\]
so
\[
	\dim_k (I'/I)_2=j.
\]
This implies any set of
generators for $I$ must have at least $j$ elements. Since $f_1,
\ldots, f_j$ is such a set, it is minimal.
\end{proof}

@@

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add_pts}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan'$ be $\mathbb{Q}$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that minimally generate\todo{check minimal generation actually works} $R_\halfcan$ over $\halfcan$' and initial terms of relations that minimally generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Grobner Basis).  The proofs of these Lemmas all follow similar patterns, so we will present a few \todo{give the precise number we show} and leave the rest to the reader.

@@

In particular, if $\frac{\alpha}{\beta}\le 1$, then $I$ is minimally generated by elements with leading terms of the form $y_i x_j$ and $y_i y_j$.
%\item If $\gin_\prec(I')=\initial_\prec(I')$ then $\gin_\prec(I)=\initial_\prec(I)$.
\item[(c)] A Grobner basis with the initial terms $y_i x_j$, $y_i y_j$,
	and $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
	\frac{\alpha}{\beta} \rfloor} + 1}$ (really the subset of these with positive subscripts) as shown above minimally
generates $I'$.
\end{enumerate}
\end{lem}

@@

\begin{lem}\label{lem:relations_from_generators_induction} 
Suppose $\halfcan$ and $\halfcan'$ are $\mathbf{Q}$ divisors on a curve $X$ with $D=D'+\frac{\alpha}{\beta}P$.  Suppose $R_{\halfcan'}$ is generated by $x_1, \ldots, x_m$ and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over $R_{\halfcan'}$; specifically, suppose $S$ is a $k$ basis for $R_\halfcan$ over $R_{\halfcan'}$ such that $S$ consists only of monomials in $x_1, \ldots, x_m, y_1, \ldots, y_n$ and each $m\in S$ is not contained in $R_{\halfcan'}$.  Let $T\subseteq R_\halfcan - (S \cup R_{\halfcan'})$, such that any monomial of $(k[x_1, \ldots, x_m, y_1, \ldots, y_n])_{\ge 2}$ (the degrees at least 2 with respect to the regrading of each $x_1, \ldots, x_m, y_1, \ldots, y_n$) is either in $S$, $\langle T\rangle$, or $R_{\halfcan'}$.  Moreover, suppose there is some ordering $\prec$ such that if $f\in S$ and $g\not\in S, g\not\in R_{\halfcan'}$ is a quadratic monomial in $x_1, \ldots, x_m, y_1, \ldots, y_n$ in the same degree as $f$, then $g\prec f$.  Then if $I$ and $I'$ are the ideal of relations of $R_{\halfcan'}$ and $R_{\halfcan}$ with the given generating sets, then 
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]\\
		     & + \langle T \rangle.
\end{align*}
Furthermore, if $T$ consists only of quadratic monomials, then $I$ is minimally generated over $I'$ by a Grobner Basis with an initial term for each element of $T$.
\end{lem}
\begin{proof}

@@

In the case when all $\frac{\alpha}{\beta}\le 1$, Lemma
~\ref{lem:minimal_quadratic} immediately tells us that the Grobner
basis defined above is minimal.
\todo{What about $\frac{\alpha}{\beta} > 1$?}
\end{proof}

@@

%\item A Grobner basis with initial terms described above minimally generates $I'$ over $I$.

@@

\ssec{Genus One Exceptional Cases}
\label{ssec:g - 1-exceptional}
\todo{remove stuff about tame and separably rooted, change to $\BQ$ divisor.}
Let $\sx$ be a tame, seperably rooted stacky curve of genus 1. Let $P,Q$ are distinct hyperelliptic fi xed points, and $R$ be a point on $\sx$.

@@

\begin{center}
\label{table:g - 1-exceptional}
\textbf{Table (IV)}: Genus 1 exceptional cases, where $L' = P - Q + \frac{e- 1}{2e}R$, with $P \neq Q$.
\end{center}
\iffalse
\begin{rem}
\label{rem:g - 1-exceptional-e}
Note that for $e > 5$ generators and relations for $L = P - Q + \frac{e- 1}{2e}R$, which has signature $(1;e;0)$, can be computed inductively $L' = P - Q + \frac{(e-2) - 1}{2(e - 2)}R$ using 
Theorem ~\ref{thm:ram_order_ind}.
\end{rem}
\fi

We now elaborate on the computation of the results in 
Table ~\ref{table:g - 1-exceptional}. We have already explained the case of 
$L' = P - Q + \frac{1}{5}P_1$ above in Example ~\ref{eg:exception- 1-5}. The other cases are similar.

\iffalse
\begin{example}
\label{eg:exceptional- 1-0}
In the case $L=P - Q.$ Then, we have $R_{(2k+ 1)L} \cong 0$ while $R_{2kL} \cong k$, with a single generator in degree 2. Therefore, $R_L \cong k[y]$, where $y$ is a generator in degree 2. 
\end{example}

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $\max(5, e_1, \ldots, e_r)$ and has minimal relations in degree at most $2e,$ so long as $\sigma$ does not lie in a
finite list of exceptional cases, as listed in Table
~\ref{table:g - 1-exceptional}.
\end{thm}

@@

Note that we are assuming $L$ is not one of the exceptional cases listed in Table ~\ref{table:g - 1-exceptional}, so we may either assume $\sx$ has 1 stacky point with $e_1 > 5$ or at least 2 stacky points

>>>>

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$ then 
$R(\sx,\Delta, \halfcan)$ has minimal relations in degree at most $\max(11, 2e_1,
\ldots, 2e_r).$ Otherwise, $R(\sx, \Delta, \halfcan)$ has minimal relations in degree at most
$\max(10, 2e_1, \ldots, 2e_r),$ so long as $\sigma$ does not lie in a finite list of
exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g_0_exceptional}
for signatures with $g = 0$.
\end{thm}

@@

Note that in the course of the proof of Theorem
~\ref{thm:g_0_generators_relations} we will not merely obtain
bounds on the degrees of the generators and relations. In fact, if
the genus $g = 0$ or $g = 1$, the proof yields an explicit system
of generators and the initial ideal of relations. While in the
genus $g \geq 2$ case, the proof does not always yield an explicit
system of generators and relations, if the halfcanonical divisor
$\halfcan$ can be written as $\halfcan' + \Lambda$ with $(\halfcan',
0, X)$ a log spin curve, and a system of generators and relations for
$ \halfcan'$ is known, then our proof provides an inductive
procedure for determining the generators and initial ideal of
relations for $\halfcan$. In fact, such a system of generators and
relations are known for $\halfcan'$ in many cases, as detailed in
Neves ~\cite[ Section III.4]{neves:halfcan}.

@@

\section{Background}
\label{sec:background}
\todo{Reference independence of base field}
Look at Spencer-Kodaira \cite{kodaira:complex-manifolds}.
Include the following:
\begin{enumerate}
	\item Note that this result is independence of base field
	\item background on Gr\"{o}bner bases
	\item Notation $h^0 = \dim H^0$.
	\item Why we have these fractions $\frac{e_i - 1}{2e_i}$ and why $e_i$ must be odd.
\end{enumerate}

@@

\noindent
determines map of rings, no longer necessarily graded. 
By assumption, $I'$ is generated over $I$ as a $k$-vector space by elements $f_1, \ldots, f_n$ whose initial terms have degree 2 and are linearly independent. By the theory of Gr\"{o}bner basis,
\[
	\dim_k(B'/B)_2=\dim_k(k[x_1, \ldots, x_n, y_1, \ldots, y_m]/k[x_1, \ldots, x_n])_2 - j
\]
so
\[
	\dim_k (I'/I)_2=j.
\]
This implies any set of
generators for $I$ must have at least $j$ elements. Since $f_1,
\ldots, f_j$ is such a set, it is minimal.
\end{proof}

@@

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add_pts}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan'$ be $\mathbb{Q}$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that minimally generate\todo{check minimal generation actually works} $R_\halfcan$ over $\halfcan$' and initial terms of relations that minimally generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Gr\"{o}bner Basis).  The proofs of these Lemmas all follow similar patterns, so we will present a few \todo{give the precise number we show} and leave the rest to the reader.

@@

In particular, if $\frac{\alpha}{\beta}\le 1$, then $I$ is minimally generated by elements with leading terms of the form $y_i x_j$ and $y_i y_j$.
%\item If $\gin_\prec(I')=\initial_\prec(I')$ then $\gin_\prec(I)=\initial_\prec(I)$.
\item[(c)] A Gr\"{o}bner basis with the initial terms $y_i x_j$, $y_i y_j$,
	and $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
	\frac{\alpha}{\beta} \rfloor} + 1}$ (really the subset of these with positive subscripts) as shown above minimally
generates $I'$.
\end{enumerate}
\end{lem}

@@

\begin{lem}\label{lem:relations_from_generators_induction} 
Suppose $\halfcan$ and $\halfcan'$ are $\mathbf{Q}$ divisors on a curve $X$ with $D=D'+\frac{\alpha}{\beta}P$.  Suppose $R_{\halfcan'}$ is generated by $x_1, \ldots, x_m$ and $R_{\halfcan}$ is minimally generated by $y_1, \ldots, y_n$ over $R_{\halfcan'}$; specifically, suppose $S$ is a $k$ basis for $R_\halfcan$ over $R_{\halfcan'}$ such that $S$ consists only of monomials in $x_1, \ldots, x_m, y_1, \ldots, y_n$ and each $m\in S$ is not contained in $R_{\halfcan'}$.  Let $T\subseteq R_\halfcan - (S \cup R_{\halfcan'})$, such that any monomial of $(k[x_1, \ldots, x_m, y_1, \ldots, y_n])_{\ge 2}$ (the degrees at least 2 with respect to the regrading of each $x_1, \ldots, x_m, y_1, \ldots, y_n$) is either in $S$, $\langle T\rangle$, or $R_{\halfcan'}$.  Moreover, suppose there is some ordering $\prec$ such that if $f\in S$ and $g\not\in S, g\not\in R_{\halfcan'}$ is a quadratic monomial in $x_1, \ldots, x_m, y_1, \ldots, y_n$ in the same degree as $f$, then $g\prec f$.  Then if $I$ and $I'$ are the ideal of relations of $R_{\halfcan'}$ and $R_{\halfcan}$ with the given generating sets, then 
\begin{align*}
	\initial(I) & = \initial(I') k[x_1, \ldots, x_m, y_1, \ldots, y_n]\\
		     & + \langle T \rangle.
\end{align*}
Furthermore, if $T$ consists only of quadratic monomials, then $I$ is minimally generated over $I'$ by a Gr\"{o}bner Basis with an initial term for each element of $T$.
\end{lem}
\begin{proof}

@@

In the case when all $\frac{\alpha}{\beta}\le 1$, Lemma
~\ref{lem:minimal_quadratic} immediately tells us that the Gr\"{o}bner
basis defined above is minimal.
\todo{What about $\frac{\alpha}{\beta} > 1$?}
\end{proof}

@@

%\item A Gr\"{o}bner basis with initial terms described above minimally generates $I'$ over $I$.

@@

\ssec{Genus One Exceptional Cases}
\label{ssec:g-1-exceptional}
\todo{remove stuff about tame and separably rooted, change to $\BQ$ divisor.}
Let $\sx$ be a tame, seperably rooted stacky curve of genus 1. Let $P,Q$ are distinct hyperelliptic fi xed points, and $R$ be a point on $\sx$.

@@

\begin{center}
\label{table:g-1-exceptional}
\textbf{Table (IV)}: Genus 1 exceptional cases, where $L' = P - Q + \frac{e- 1}{2e}R$, with $P \neq Q$.
\end{center}
\iffalse
\begin{rem}
\label{rem:g-1-exceptional-e}
Note that for $e > 5$ generators and relations for $L = P - Q + \frac{e- 1}{2e}R$, which has signature $(1;e;0)$, can be computed inductively $L' = P - Q + \frac{(e-2) - 1}{2(e - 2)}R$ using 
Theorem ~\ref{thm:ram_order_ind}.
\end{rem}
\fi

We now elaborate on the computation of the results in 
Table ~\ref{table:g-1-exceptional}. We have already explained the case of 
$L' = P - Q + \frac{1}{5}P_1$ above in Example ~\ref{eg:exception- 1-5}. The other cases are similar.

\iffalse
\begin{example}
\label{eg:exceptional-1-0}
In the case $L=P - Q.$ Then, we have $R_{(2k+ 1)L} \cong 0$ while $R_{2kL} \cong k$, with a single generator in degree 2. Therefore, $R_L \cong k[y]$, where $y$ is a generator in degree 2. 
\end{example}

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $\max(5, e_1, \ldots, e_r)$ and has minimal relations in degree at most $2e,$ so long as $\sigma$ does not lie in a
finite list of exceptional cases, as listed in Table
~\ref{table:g-1-exceptional}.
\end{thm}

@@

Note that we are assuming $L$ is not one of the exceptional cases listed in Table ~\ref{table:g-1-exceptional}, so we may either assume $\sx$ has 1 stacky point with $e_1 > 5$ or at least 2 stacky points
